---
title: "Covid91 vaccine study FinalFall2022"
author: "Elizabeth Rivera"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
malesdata=filter(FinalData,sex=="M")
```

$H_0$: The treatment has no effect on males.  
$H_A$: The treatment has a significant effect on males. 


### Graphical Description  
```{r}
barchartGC(~infected + treatment,data=malesdata)
barchartGC(~infected + treatment,data=malesdata, type="percent")

```


Based on the graph above, among the men who contracted COVID91 about twice as many received the placebo as received the COVID91 treatment. From this graph it looks like the treatment is effective for males.  


### Numerical Description  

```{r}
table1 <- xtabs(~infected + treatment, data=malesdata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical data, among the males that contracted COVID91 68.44% received the placebo, and 31.56% of males received the treatment. The data demonstrates the treatment is effective for males.  

### Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi Squared test has a p-value of 0.000000000005307 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on males.

The Fisher's Test has a p-value of 0.000000000002364 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on males. 

Males who received the placebo are 2.18 times more likely to contract COVID91 than males who received the treatment.    

### Discussion and Conclusion

Based on the data above, among the males who contracted COVID91 68.44% received the placebo and 31.56% received the treatment. According to the data, among the males who received the placebo, the odds are that they are 2.18 times more likely to contract COVID91 than males who received the treatment.  


## Females

```{r}
femalesdata=filter(FinalData,sex=="F")
```

$H_0$: The treatment has no effect on females.  
$H_A$: The treatment has a significant effect on females.   

### Graphical Description  
```{r}
barchartGC(~infected + treatment,data=femalesdata)
barchartGC(~infected + treatment,data=femalesdata, type="percent")

```

Based on the graph above, among the females who contracted COVID91 about 44% received the placebo compared to 56% received the COVID91 treatment. Among females who received either the placebo or treatment, but did not contract COVID91 were very similar at about 50%. From this graph it looks like the treatment is not very effective for females.  

### Numerical Description
```{r}
table1 <- xtabs(~infected + treatment, data=femalesdata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical data, among the females that contracted COVID91 42.82% received the placebo, and 57.18% of females received the treatment. The data demonstrates the treatment is not effective for females.  

### Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi Squared test has a p-value of 0.005747 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on females.  

The Fisher's Test has a p-value of 0.004887 which is less than 0.05 meaning we reject the null hypothesis.  

We conclude that the treatment does have an effect on females.Females who received the placebo are 0.74 times more likely to contract Covid91 than females who received the treatment.

### Discussion and Conclusion    

Based on the data above, among the females who contracted COVID91 42.82% received the placebo and 57.18% received the treatment. According to the data, among the females who received the placebo, the odds are that they are 0.74 times more likely to contract COVID91 than females who received the treatment.  


## LGBTQ
```{r}
lgbtqdata=filter(FinalData,LGBTQ==LGBTQ)
```

$H_0$: The treatment has no effect on LGBTQ community.  
$H_A$: The treatment has a significant effect on LGBTQ community.  

### Graphical Description  
```{r}
barchartGC(~infected + treatment,data=lgbtqdata)
barchartGC(~infected + treatment,data=lgbtqdata, type="percent")

```
  
Based on the graph above, among the females who contracted COVID91 about 44% received the placebo compared to 56% received the COVID91 treatment. Among the LGBTQ community who received either the placebo or treatment, but did not contract COVID91 were very similar at about 50%. From this graph it looks like the treatment is not very effective for the LGBTQ community.

```{r}
table1 <- xtabs(~infected + treatment, data=lgbtqdata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical data, among the LGBTQ community that contracted COVID91 55.56% received the placebo, and 44.44% of the LGBTQ community received the treatment. The data demonstrates the treatment is not effective for the LGBTQ community.   

### Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
Chi Squared test has a p-value of 0.004136 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on the LGBTQ community.  

The Fisher's Test has a p-value of 0.003736 which is less than 0.05 meaning we reject the null hypothesis.  

We conclude that the treatment does have an effect on the LGBTQ community. Individuals in the LGBTQ community who received the placebo are 1.25 times more likely to contract Covid91 than individuals in the LGBTQ community who received the treatment.

### Discussion and Conclusion    

Based on the data above, among the individuals of the LGBTQ community who contracted COVID91 55.56% received the placebo and 44.44% received the treatment. According to the data, among the LGBTQ community who received the placebo, the odds are that they are 1.25 times more likely to contract COVID91 than individuals in the LGBTQ community who received the treatment.      

## Druggies
```{r}
druggiesdata=filter(FinalData,DrugUser=="yes")
```

$H_0$: The treatment has no effect on drug users   
$H_A$: The treatment has a significant effect on drug users.  

##Graphical Description  
```{r}
barchartGC(~infected + treatment,data=druggiesdata)
barchartGC(~infected + treatment,data=druggiesdata, type="percent")

```  

Based on the graph above, among the drug users who contracted Covid91 about 60% more of them received the Covid91 treatment as received the placebo. The graph suggest that the treatment is not effective for drug users.  

### Numerical Description  

```{r}
table1 <- xtabs(~infected + treatment, data=druggiesdata)
rowPerc(table1)
colPerc(table1)
```  

According to the numerical data, among the drug users that contracted Covid91 16.67% of them received the placebo, and 83.33% of them received the  treatment. The data demonstrates that the treatment is not effective for drug users.   

### Inferential Results  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  

Chi Squared test has a p-value of 0.000001419 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on drug users.
The Fisher's Test has a p-value of 0.0000003119 which is less than 0.05 meaning we reject the null hypothesis. We conclude that the treatment does have an effect on drug users.  
Drug users who received the placebo are 0.21 times as likely to contract Covid91 than the drug users who received the treatment, meaning the treatment has a negative effect on drug users.  


### Discussion and Conclusion  
Based on the data above, among the drug users who contracted Covid91 16.67% of drug users received the placebo and 83.33% of drug users received the treatment. According to the data, among the drug users who received the placebo, the odds are that they are 0.21 times more likely to contract Covid91 than the drug users who received the treatment.Therefore we reject the null hypothesis since the treatment has a significant negative effect on drug users since it makes them more susceptible to contracting Covid 91.  








# Overall Results and Conclusions